Fractional Types

### 2019-11-20 fracGC/p.tex
reorder sections

\paragraph*{Outline.} The paper solves the ancilla problem in
reversible and quantum computation using a novel concept:
\emph{fractional types}. In the next section, we introduce the problem
of ancilla, motivate its importance, and explain the limitations of
current solutions. In Sec.~\ref{sec:pi}, we review existing work by
James et al.~\cite{rc2011,rc2012,James:2012:IE:2103656.2103667} that
introduces a reversible programming language built using isomorphism
of types. Although it is conceivable that the concept of fractional
types would apply to general-purpose languages (after some
adaptation), its natural technical definition is using symmetries
present in categorical model of type isomorphisms. We present such a
definition in Sec.~\ref{sec:dyn}: that definition captures the main
ideas of fractional types and partially solves a more general and more
expressive instance of the ancilla problem. The approach however
shares the limitation of current solutions in that it relies on
runtime checks for programmer's assertions that values have been
restored to their initial value before de-allocation. To remedy this
limitation, we refine the type system in Sec.~\ref{sec:dep} to keep
track of values in focus. Technically this requires the introduction
of pointed types on top of which are defined both a monad and co-monad
of singleton types. To justify the claim that fractional types solve
the management of ancilla-based scratch space, we present in
Sec.~\ref{sec:space} an operational semantics that exposes memory
management details using a realization of fractional types as
first-class dependently-typed garbage collectors. In
Sec.~\ref{sec:cat} we present a rich collection of examples
illustrating the expressive power of fractional types. We conclude
with a discussion of related work and a summary of our results.

@@

\subsection{Extraction}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Trace}
\label{sec:cat}

>>>>

\paragraph*{Outline.} The paper solves the ancilla problem in
reversible and quantum computation using a novel concept:
\emph{fractional types}. In the next section, we introduce the problem
of ancilla, motivate its importance, and explain the limitations of
current solutions. In Sec.~\ref{sec:pi}, we review existing work by
James et al.~\cite{rc2011,rc2012,James:2012:IE:2103656.2103667} that
introduces a reversible programming language built using isomorphism
of types. Although it is conceivable that the concept of fractional
types would apply to general-purpose languages (after some
adaptation), its natural technical definition is using symmetries
present in categorical model of type isomorphisms. We present such a
definition in Sec.~\ref{sec:dyn}: that definition captures the main
ideas of fractional types and partially solves a more general and more
expressive instance of the ancilla problem. To make the memory
management explicit, Sec.~\ref{sec:space} provides an operational
semantics that exposes memory management details using a realization
of fractional types as first-class garbage collectors. In
Sec.~\ref{sec:dep} we refine the type system to keep track of ancillae
values. Technically this requires the introduction of pointed types on
top of which are defined both a monad and co-monad of singleton
types. We conclude with a discussion of related work and a summary of
our results.

@@

%%%%%%%%%
\subsection{Extraction}

@@

%%%%%%%%%%%
\subsection{Examples}
\label{sec:cat}
